---
title: "GitHub for Epidemiologists"
author: "Mark Klose"
date: "`r Sys.Date()`"

output: 
  html_document:
    theme: journal
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Why Source Control?

So we can stalk each other's work! 

In all seriousness, source control like GitHub provides a place to store backups and share code easily within a team or to the public. My intention is to write research articles (and hopefully my dissertation) using source control tools such as Git. That's if they don't kick me out first. 

Some negatives about source control: although Git and GitHub are two dominant source control tools, they are not the only ones. Depending on University/Corporate affiliations, you may have to "fit into" another predefined system. In addition, if you've never used the command line, this can be somewhat daunting. 

Some positives about source control: you get to use the command line and look super cool and impressive, especially after you've looked at the other tutorial on [Terminal for Epidemiologists](fe_terminal.html). 


# Step -1: Installing Git {.tabset}

To use Git, we need to have Git installed first. A good way to check if you have Git already installed is to open a terminal environment and type:

```
git --version 
```

which could produce output like:
```{r version_ex, echo=FALSE}
cat(system("git --version", intern=TRUE))
```



If this does NOT work for you, there are options to download Git:

## Manual Download
To manually download Git, you can navigate to the [Git Download Site](https://git-scm.com) to download the latest version of Git. Follow the instructions on the downloader to install Git. 

Then go install Homebrew or Chocolatey, because they are super useful make your life a whole lot easier.  


## Homebrew (MacOS)
So, to use Homebrew, you [need to have homebrew installed on your computer](https://brew.sh) (and note this is only for MacOS systems). 

Then, to install Git, you just need to run the following command in the terminal:

```
brew install git
```


## Chocolatey (Windows)
So, to use Chocolatey, you [need to have chocolatey installed on your computer](https://chocolatey.org/install) (and note this is only for Windows Users). 

Then, to install Git, you just need to run the following command in the terminal:

```
choco install git
```

## Linux Distributions
If you're using Linux, then you probably have a distro specific package manager. My suggestion? Google how to install Git with that package manager. You're using Linux, you probably have a good idea about what you are doing. 

Everything is some variation of:
```
<package manager> install git
```


# Step 0: Make a Repository


# Step 1: Clone Repository
 ```
 git clone <URL from GitHub>
 ```

# Step 2: ```.gitignore``` and more! {.tabset}
A ```.gitignore``` file allows Git to ignore any unnecessary files that fit a certain pattern. Before running any Git command, the system looks for this file and disregards any matches. For example, the MacOS system sometimes generates a ```.DS_Store``` file, RStudio generates an ```.RData``` file, maybe you keep a set of test credentials in a ```test.credentials``` file, and if they are included in the ```.gitignore``` file, then they would not be added to the staging area (which you will learn about soon!). 

To create a ```.gitignore``` file, there are two options. The first is that upon creating a repository, GitHub includes a checkbox to make a file (and includes various templates depending on which programming language is being used). The second option is to manually make the file.

## MacOS
To create files from the terminal, perform the following command:
```
touch .gitignore
```
where ```touch``` tells the operating system to create the file if it does not already exist, otherwise, update the last modified time. 


Then, to open the file from the command line, perform the following command:
```
oepn .gitignore
```
which opens the file in the system default editor. 


## Windows
To create files from the terminal, perform the following command:
```
touch .gitignore
```
where ```touch``` tells the operating system to create the file if it does not already exist, otherwise, update the last modified time. 


Then, to open the file from the command line, perform the following command:
```
explorer .gitignore
```
which opens the file in the system default editor. 

# {.unlisted .unnumbered}
Within the ```.gitignore``` file, you have a couple of options for what to ignore. 

## Ignore a single file
Within the ```.gitignore``` file, all you need to add is the filename to ignore it. 

```
> cat .gitignore
mark.txt
names/name1.txt
```
would ignore the file called ```mark.txt``` and the file ```name1.txt``` within the ```names``` directory. 

## Ignore a directory of files
Within the ```.gitignore``` file, all you need to add is the directory of files to ignore it. 

```
> cat .gitignore
names/
```
would ignore all files within the ```names``` directory, like if ```name1.txt```, ```name2.txt```, and ```name3.txt``` were all present, they would not be added to the staging area. 

## Ignore a file that fits a certain pattern
Within the ```.gitignore``` file, all you need to add is the add the pattern of files to ignore them. 
```
> cat .gitignore
*.knit.md
```
would ignore all intermediary files produced by RMarkdown. 

# Step 3: Add Files to Staging/Tracking Area
After changes have been made to files, you need to ultimately ```commit``` them to your repository in order to save the changes "permanently". The first step in this process is to add files that you want to commit to the staging area. 

There are two most commonly used commands for ```add```:
```
git add .
```
adds all files that have been edited AND do not fit the patterns of the ```.gitignore``` file. 

and 
```
git add <filename>
```
adds a specific file to the staging area. This is especially useful when first committing a ```.gitignore``` file.

Remember that at this point, you have files that you WANT to commit to your repository, but they are not committed just yet. 


# Step 4: Commit to (Local) Repository
After adding all files you want to commit to your repository, you ```commit``` them to the repository in order to make a permanent snapshot. The files that are ```commit```ted are all the ones in the staging area.
 
 To perform a ```commit```, use the following command:
 ```
 git commit -m "Put a meaningful message here"
 ```
 The ```-m``` command says "I am including a message with this commit during this step". Otherwise, during the `push` process, it will force you to enter a message using a text editor called Vim (which can be very confusing if you've never used it before). 


# Step 5: Push to (Remote) Repository

## Common Issues: {.tabset}
This is most likely where problems will arise. Some of the most common errors are listed below. 

### Non-fast forward
Someone else who has access to the remote repository has made a change before you could; you need your repository to look like theirs first. To do this, perform:

```
git pull
```

If the message says something about `merge` errors, then check the "Merge Errors" tab. 

### A scary text editor popped up!
This is most likely from missing the `-m` flag on the ```git commit -m``` step. This text editor is called Vim, and it is 100% navigable by keyboard.

To insert a commit message, press `i` (for insert), then type the message. Once finished, press `ESC` key to get to the Vim command line, then type `:wq` to save and quit. This should solve most problems. 

### Merge Errors
When pulling data from a remote repository, it is possible that your local code and the remote code are different enough that `git` does not know how to handle the differences. 


# Step 6: Repeat Process
After pushing to the remote repository, edit files, `add` them to the staging area, `commit` them to the local repository, then `push` them to the remote repository (with maybe a `pull` operation in between)! If you get stuck, it never hurts to check your ```git status``` to see what files are or are not staged. 


