---
title: "GitHub for Epidemiologists"
author: "Mark Klose"
date: "`r Sys.Date()`"

output: 
  html_document:
    theme: journal
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Why Source Control?

So we can stalk each other's work! 

In all seriousness, source control like GitHub provides a place to store backups and share code easily within a team or to the public. My intention is to write research articles (and hopefully my dissertation) using source control tools such as Git. That's if they don't kick me out first. 

Some negatives about source control: although Git and GitHub are two dominant source control tools, they are not the only ones. Depending on University/Corporate affiliations, you may have to "fit into" another predefined system. In addition, if you've never used the command line, this can be somewhat daunting. 

Some positives about source control: you get to use the command line and look super cool and impressive, especially after you've looked at the other tutorial on [Terminal for Epidemiologists](fe_terminal.html). 


# Step -1: Installing Git {.tabset}

To use Git, we need to have Git installed first. A good way to check if you have Git already installed is to open a terminal environment and type:

```
git --version 
```

which could produce output like:
```{r version_ex, echo=FALSE}
cat(system("git --version", intern=TRUE))
```



If this does NOT work for you, there are options to download Git:

## Manual Download
To manually download Git, you can navigate to the [Git Download Site](https://git-scm.com) to download the latest version of Git. Follow the instructions on the downloader to install Git. 

Then go install Homebrew or Chocolatey, because they are super useful make your life a whole lot easier.  


## Homebrew (MacOS)
So, to use Homebrew, you [need to have homebrew installed on your computer](https://brew.sh) (and note this is only for MacOS systems). 

Then, to install Git, you just need to run the following command in the terminal:

```
brew install git
```


## Chocolatey (Windows)
So, to use Chocolatey, you [need to have chocolatey installed on your computer](https://chocolatey.org/install) (and note this is only for Windows Users). 


## Linux Distributions
If you're using Linux, then you probably have a distro specific package manager. My suggestion? Google how to install Git with that package manager. You're using Linux, you probably have a good idea about what you are doing. 



# Step 0: Make a Repository


# Step 1: Clone Repository
 

# Step 2: ```.gitignore``` and more!


# Step 3: Add Files to Staging/Tracking Area


# Step 4: Commit to (Local) Repository


# Step 5: Push to (Remote) Repository


# Step 6: Repeat Process



