---
title: "R Generated Resume"
author: Mark Klose
date: "`r Sys.Date()`"
output:
  pagedown::html_resume:
    css:
      - R_Resume_Maker_files/override.css
      - resume
    # set it to true for a self-contained HTML page but it'll take longer to render
    self_contained: false
---

```{r, include=FALSE}
# Basic knitr requirements
knitr::opts_chunk$set(
  results='asis', 
  echo = FALSE
)

# Import libraries
libs <- c("tidyverse", "glue", "stringr")
# Install libraries if needed
lapply(libs, function(x){
  # if(!require(x)) install.packages(x, repos = "http://cran.us.r-project.org")
  library(x, character.only = TRUE)
})



# Load R Scripts to
resume_file <- "./details.csv"
# Read in Resume Data
source("gather_data.R")
source("make_sections.R")

```


Aside
================================================================================

```{r}
make_contact(info[["contact"]])
```

Skills {#skills}
--------------------------------------------------------------------------------

```{r}
make_skills(info[["skills"]])
```


Disclaimer {#disclaimer}
--------------------------------------------------------------------------------

This resume was made with the R package [**pagedown**](https://github.com/rstudio/pagedown).
Last updated on `r Sys.Date()`.



Main
================================================================================

```{r}
make_blurb(info[["blurb"]], priority = 0)
make_education(info[["education"]])
make_research(info[["research"]], priority = 0)
make_prof(info[["work"]], priority=0)
```


Teaching Experience {data-icon=chalkboard-teacher}
--------------------------------------------------------------------------------

### Introduction to R Language for Beginners.

Instructor of R and Data Mining Training Courses at SupStat Inc.

Beijing, China

2014

### Computational Biology and Bioinformatics.

Teaching assistant of GBS CB2-201 courses at UAB

AL, USA

2016 - 2017


Selected Publications and Posters {data-icon=file}
--------------------------------------------------------------------------------

### Genetic and epigenetic signals are found predictive to the distribution of intra-individual divergence of alternative splicing.

Poster for 2013 International Conference of Genomics

Qingdao, China

2014

**Yu L**, Chen B, Zhang Z.

### ESCRT-0 complex modulates Rbf mutant cell survival by regulating Rhomboid endosomal trafficking and EGFR signaling.

J Cell Sci. 2016 May 15;129(10):2075-84.

N/A

2016 

Sheng Z, **Yu L**, Zhang T, Pei X, Li X, Zhang Z and Du W. 
